Collaborative Spatial Reuse in Wireless Networks via Selfish Multi-Armed Bandits

### 2018-11-13 Article - LaTeX files/collaborative_spatial_reuse_via_mabs.tex
Added authors biographies and photos.
Version submitted on 13th, November 2018. Includes the new version of the manuscript and the response letter.

\usepackage{color}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{algorithmic}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\algsetup{linenosize=\small}
\usepackage[table,xcdraw]{xcolor}
\usepackage{multirow}
\usepackage[super]{nth}
\usepackage{graphicx,adjustbox}
\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\usepackage{comment}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{xspace}
\usepackage{siunitx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{soul}
\usepackage{url}
\usepackage{tablefootnote}
\DeclareMathOperator{\E}{\mathbb{E}} % Expectation Symbol
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{footnote}
\usepackage[misc,geometry]{ifsym} 
\makesavenoteenv{tabular}

@@

\journal{Computer Networks}

@@

\author[label1, label4]{Francesc Wilhelmi}\ead{francisco.wilhelmi@upf.edu}
\author[label2]{Cristina~Cano}
\author[label3]{Gergely~Neu}
\author[label1]{Boris~Bellalta} 
\author[label3]{Anders~Jonsson}
\author[label1]{Sergio~Barrachina-Mu\~noz}
\address[label1]{Wireless Networking Research Group (WN-UPF), 08002 Barcelona, Spain}
\address[label2]{Wireless Networks Research Group (WINE-UOC), 08860 Castelldefels (Barelona), Spain}
\address[label3]{Artificial Intelligence and Machine Learning Research Group (AIML-UPF), 08002 Barcelona, Spain}

\begin{abstract}
Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to enable Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners and WNs operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action.
\end{abstract}

\begin{keyword}
High-Density Wireless Networks, Interference, Decentralized learning, Multi Armed Bandits, Spatial Reuse, Resource Allocation.
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  I. INTRODUCTION 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
Due to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) sharing the same channel resources. Due to the decentralized nature of such kind of deployments, the lack of organization and/or agreement on sharing policies leads to inefficient resources allocation, both in spectral and spatial domains. An illustrative example about the inefficiency noticed in WNs regarding channel sharing can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and of the channel occupancy \cite{akella2007self}. In decentralized environments, providing a model to obtain the optimal solution can be extremely complex. In WNs, the interactions between devices depend on many features (such as position, environment, transmit power) and are hard to derive. On top of this, network dynamics, in terms of user arrivals and departures, makes the resource allocation problem even harder.

In order to address the SR problem in decentralized WNs, we focus attention on Reinforcement Learning (RL), which has recently become very popular to solve many well-known problems in wireless communications. Some examples can be found for packet routing \cite{littman1993distributed}, Access Point (AP) selection \cite{bojovic2011supervised, bojovic2012neural}, optimal rate sampling \cite{combes2014optimal}, or energy harvesting in heterogeneous networks \cite{miozzo2015distributed}. Using online learning, a learner (or agent) obtains data in a sequential order and uses it to predict future good-performing actions. Online learning is, therefore, useful to cope with complex and dynamic environments. This background encourages us to approach a solution for the decentralized SR problem in WNs through online learning techniques. In particular, we exploit dynamic channel selection and transmit power adjustment for dealing with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness.

From the family of online algorithms, we propose the utilization of Multi-Armed Bandits (MABs) \cite{BCB12}, which is a well-known model in the online learning literature for solving resource allocation problems. In MABs, a given agent seeks to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. %(exploitation focuses on obtaining the maximum instantaneous reward, while exploration aims to maximize the long-term reward)	
While exploitation selects actions with the goal of maximizing long-term reward given the current estimate, exploration selects actions with the goal of improving the estimate. Unlike for classical RL, MABs do not consider states\footnote{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge on its current situation, an agent can define state-specific strategies that maximize its profits.} in general, which can be hard to define for the decentralized SR problem. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions between nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the system complexity, since the state not only depends on the actions taken by a given node, but also on the adversaries behavior.

This work extends our previous results presented in \cite{wilhelmi2017implications}, which were obtained by applying a stateless variation of Q-learning to the SR problem in WNs. Here we generalize the contributions done by implementing several action-selection strategies to allow WNs choose the channel and transmit power to use based on the resulting performance. In this paper, then, we provide a tutorial-based application of MABs to the decentralized SR problem, in order to maximize performance while observing the result from using different SR settings. According to that, we aim to illustrate the main benefits and drawbacks of applying different well-known learning techniques in a selfish way, i.e., independent WNs base their strategy on their own experienced performance. On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, in order to shed some light on the future of decentralized approaches. Notably, we observe that even though players act selfishly, there are algorithms %learn to play collaborative actions
learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments as we get rid of the need for message passing, or inference of neighboring conditions.

The main contributions of this work are summarized below:
\begin{itemize}		
	\item We devise the feasibility of applying MAB algorithms as defined in the online learning literature to solve the resource allocation problem in WNs.
	\item We study the impact of different parameters intrinsic to the action-selection strategies considered (e.g., exploration coefficients, learning rates) on network performance.
	\item We show that there are algorithms learn to play collaborative actions even though the WNs act selfishly, which is appealing to practical application in chaotic and dynamic environments. In addition, we shed some light on the root causes of this phenomena. 
\end{itemize}

The remaining of this document is structured as follows: Section \ref{section:related_work} outlines relevant related work. Section \ref{section:mabs} introduces the proposed learning algorithms and their practical implementation for the resource allocation problem in WNs. Then, Section \ref{section:system_model} presents the simulation scenarios and the considerations taken into account. The simulation results are later presented in Section \ref{section:performance_evaluation}. Finally, Section \ref{section:conclusions} provides some final remarks.

%%%%%%%%%%%%%%%%%%%%%%%
%%%  II. RELATED WORK 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{section:related_work} 		
In this paper we approach the decentralized SR problem through Dynamic Channel Allocation (DCA) and Transmit Power Control (TPC). While the former aims to allocate the available frequency range among potentially overlapping WNs, the latter attempts to increase the number of parallel transmissions by adjusting the power transmitted by each WN. While DCA operates at the frequency level, TPC deals with more complex spatial interactions. For that, an approach that dynamically tunes the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality.

DCA has been extensively exploited from the centralized perspective, especially through techniques based on graph coloring \cite{riihijarvi2005frequency, mishra2005weighted}. Despite these kind of approaches allow to effectively reduce the interference between WNs, a certain degree of communication is required. Regarding decentralized methods, the authors in \cite{akl2007dynamic} propose a very simple approach in which each AP maintains an interference map of their neighbors, so that channel assignment is done through interference minimization. Unfortunately, the interactions between APs due to decentralization are not studied. Separately, \cite{chen2007improved} proposes two decentralized approaches that rely in the interference measured at both APs and STAs to calculate the best frequency channels for dynamic channel allocation. To do so, a WN, in addition to the interference sensed by its associated devices, considers other metrics such as the amount of traffic, so that some kind of coordination is required at the neighbor level (periodic reporting).

For the case of TPC, one of the first works in studying power management is \cite{elbatt2000power}, which shows that large improvements can be achieved if selecting the appropriate transmit power level in ad-hoc WNs. A centralized approach for TPC is presented in \cite{tang2014joint}, which performs power control and Rate Adaptation (RA) in subgroups of Wireless Local Area Networks (WLANs). The creation of clusters allows defining independent power levels between devices in the same group, which are useful to avoid asymmetric links. However, to represent all the possible combinations, graphs can become very large, specially in high-density deployments. Furthermore, we find a decentralized approach that relies on real-time channel measurements \cite{gandarillas2014dynamic}. The proposed mechanism (so called Dynamic Transmission Power Control) is based on a set of triggered thresholds that increase/decrease the transmit power according to the situation. The main problem is that thresholds are set empirically (based on simulations), so the mechanism may not always improve performance.

%------------------------
% Decentralised Learning: Previous work
%------------------------
\subsection{Solving WNs Resource Allocation Problems via Bandits}
\label{section:previous_work}
In the online learning literature, several MAB settings have been considered such as stochastic bandits \cite{thompson1933likelihood,lai1985asymptotically,auer2002finite}, adversarial bandits \cite{auer1995gambling,auer2002nonstochastic}, restless bandits \cite{whittle1988restless}, contextual bandits \cite{LCLS10} and linear bandits \cite{abe2003reinforcement,APS11}, and numerous exploration-exploitation strategies have been proposed such as \textit{$\varepsilon$-greedy} \cite{sutton1998reinforcement,auer2002finite}, \textit{upper confidence bound} (UCB) \cite{lai1985asymptotically,Agr95,BuKa96,auer2002finite}, \textit{exponential weight algorithm for exploration and exploitation} (EXP3) \cite{auer1995gambling,auer2002finite} and \textit{Thompson sampling} \cite{thompson1933likelihood}. The classical multi-armed bandit problem models a sequential interaction scheme between a learner and an environment. The learner sequentially selects one out of $K$ actions (often called \emph{arms} in this context) and earns some rewards determined by the chosen action and also influenced by the environment. Formally, the problem 
is defined as a repeated game where the following steps are repeated in each round $t=1,2,\dots,T$:
\begin{enumerate}
	\item The environment fixes an assignment of rewards $r_{a,t}$ for each action $a\in[K] \stackrel{\text{def}}{=} \left\{1,2,\dots,K\right\}$,
	\item the learner chooses action $a_t\in[K]$,
	\item the learner obtains and observes reward $r_{a_t,t}$.
\end{enumerate}
The bandit literature largely focuses on the perspective of the learner with the objective of coming up with learning algorithms that attempt to maximize the sum of the rewards gathered during the whole procedure (either with finite or infinite horizon). As noted above, this problem has been studied under various assumptions made on the environment and the structure of the arms. The most important basic cases are the \emph{stochastic} bandit problem where, for each particular arm $a$, the rewards are i.i.d.~realizations of random variables from a fixed (but unknown) distribution $\nu_a$, and the \emph{non-stochastic} (or \emph{adversarial}) bandit problem where the rewards are chosen arbitrarily by the environment. In both cases, the main challenge for the learner is the \emph{partial observability} of the rewards: the learner only gets to observe the reward associated with the chosen action $a_t$, but never observes the rewards realized for the other actions.

@@

To the best of our knowledge, there is very little related work on applying multi-armed bandit techniques to the problem of resource allocation in WNs. In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in LTE networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which allows to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that an optimal resource allocation exists, i.e., every BS obtains the necessary resources. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.

>>>>

\usepackage{color}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{algorithmic}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\algsetup{linenosize=\small}
\usepackage[table,xcdraw]{xcolor}
\usepackage{multirow}
\usepackage[super]{nth}
\usepackage{graphicx,adjustbox}
\usepackage{caption}
\usepackage[labelformat=simple]{subcaption}
\usepackage{comment}
\usepackage{setspace}
\usepackage{textcomp}
\usepackage{xspace}
\usepackage{siunitx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{soul}
\usepackage{url}
\usepackage{tablefootnote}
\DeclareMathOperator{\E}{\mathbb{E}} % Expectation Symbol
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{footnote}
\usepackage[misc,geometry]{ifsym} 
\makesavenoteenv{tabular}
\usepackage[utf8]{inputenc}

@@

\journal{Ad-hoc Networks}

@@

		\author[label1, label4]{Francesc Wilhelmi}\ead{francisco.wilhelmi@upf.edu}
		\author[label2]{Cristina~Cano}
		\author[label3]{Gergely~Neu}
		\author[label1]{Boris~Bellalta} 
		\author[label3]{Anders~Jonsson}
		\author[label1]{Sergio~Barrachina-Mu\~noz}
		\address[label1]{Wireless Networking Research Group (WN-UPF), 08018 Barcelona, Spain}
		\address[label2]{Wireless Networks Research Group (WINE-UOC), 08860 Castelldefels (Barelona), Spain}
		\address[label3]{Artificial Intelligence and Machine Learning Research Group (AIML-UPF), 08018 Barcelona, Spain}

		\begin{abstract}
			Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to enable Spatial Reuse (SR). \textcolor{blue}{In particular, we focus on dynamic channel selection and Transmission Power Control (TPC). We rely on Reinforcement Learning (RL), and more specifically on Multi-Armed Bandits (MABs), to allow networks to learn their best configuration}. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection, and compare their performance. \textcolor{blue}{In addition, we study the implications of selecting actions simultaneously in an adversarial setting \textcolor{blue}{(i.e., concurrently)}, \textcolor{blue}{and compare it with a sequential approach}.} Our results show that optimal proportional fairness can be achieved, even when no information about neighboring networks is available to the learners and Wireless Networks (WNs) operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, especially for $\varepsilon$-greedy and EXP3. \textcolor{blue}{These strategies, contrary to UCB and Thompson sampling, base their operation on the absolute experienced reward, rather than on its distribution.} We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. \textcolor{blue}{We also show that \textcolor{blue}{learning sequentially, even if using a selfish strategy,} contributes to minimize this variability. \textcolor{blue}{The sequential approach is therefore shown to effectively deal with the challenges posed by the adversarial settings that are typically found in decentralized WNs.}}% The latter} are parameter-free policies that perform exploration according to the reward distribution of each action.
		\end{abstract}

		\begin{keyword}
			High-Density Wireless Networks; Spatial Reuse; Resource Allocation; Decentralized learning; Multi-Armed Bandits 
		\end{keyword}

@@

	%%%%%%%%%%%%%%%%%%%%%%%
	%%%  I. INTRODUCTION 
	%%%%%%%%%%%%%%%%%%%%%%%
	\section{Introduction}
	\label{section:introduction}
	Due to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) sharing the same channel resources. \textcolor{blue}{The decentralized nature of such kind of deployments leads to a significant lack of organization and/or agreement on sharing policies. As a result, resources are typically used inefficiently. An illustrative example of this can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and the channel occupancy. Consequently, increasing the capacity of such networks has become very challenging.}

	\textcolor{blue}{Wireless networks operate in three main domains: time, frequency and space. While the first two have been largely exploited, the spatial domain still shows plenty of room for improvement. According to \cite{alawieh2009improving}, Spatial Reuse (SR) can be addressed by means of Transmission Power Control (TPC), Carrier Sense Threshold (CST) adjustment, rate adaptation (related to power control), and directional transmissions. In addition, interference cancellation can play a key role on spectral efficiency optimization \cite{miridakis2013survey}. On one side, TPC and CST adjustment aim at increasing spectral efficiency omnidirectionally. On the other hand, beamforming is meant for directional transmissions. Both beamforming and interference cancellation can be categorized as multiple antenna strategies. While the former allows to reduce the interference levels, the second one is useful to perform multiple simultaneous transmissions.}

	\textcolor{blue}{In this work, we focus on Dynamic Channel Allocation (DCA) and TPC to address the decentralized SR problem. A proper frequency planning allows to reduce the interference between wireless devices, and tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness. The application of TPC and DCA is particularly challenging by itself. The interactions among devices depend on many features (such as position, environment or transmit power) and are hard to derive. %On top of this, user arrivals and departures make the resource allocation problem even harder. 
		Including beamforming and/or interference cancellation techniques \cite{dovelos2018breaking}, on the other hand, requires first a clear understanding of TCP and DCA performance alone, and is therefore left as future work.}

	\textcolor{blue}{Motivated by these challenges,} we focus attention on Reinforcement Learning (RL), which has recently emerged as a very popular method to solve many well-known problems in wireless communications. \textcolor{blue}{RL allows to reduce the complexity generated in wireless environments by finding practical solutions. By applying RL, optimal (or near-to-optimal) solutions can be obtained without having a full understanding on the problem in advance. So, one of the main goals of this paper is to show its feasibility for the decentralized SR problem.} Some RL-based applications can be found for packet routing \cite{littman1993distributed}, Access Point (AP) selection \cite{bojovic2011supervised, bojovic2012neural}, optimal rate sampling \cite{combes2014optimal}, or energy harvesting in heterogeneous networks \cite{miozzo2015distributed}. All these applications make use of online learning, where a learner (or agent) obtains data periodically and uses it to predict future good-performing actions. Online learning is particularly useful to cope with complex and dynamic environments. This background encourages us to approach a solution for the decentralized SR problem in WNs through online learning techniques.

	From the family of online algorithms, we \textcolor{blue}{are interested on analyzing the performance of} Multi-Armed Bandits (MABs) \cite{BCB12} \textcolor{blue}{when applied to WNs. The MAB model is well-known} in the online learning literature for solving resource allocation problems. In MABs, a given agent seeks to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. Exploitation is meant to maximize the long-term reward given the current estimate, and exploration aims to improve the estimate. Unlike classical RL, MABs do not consider states\footnote{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge of its current situation, an agent can define state-specific strategies that maximize its profits.} in general, which can be hard to define for the decentralized SR problem presented in this work. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions among nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the system complexity. Therefore, the obtained reward does not only depends on the actions taken by a given node, but also on the adversaries behavior.

	\textcolor{blue}{This article extends our previous results presented in \cite{wilhelmi2017implications}. Here we generalize the contributions done by implementing several action-selection strategies to find the best combination of frequency channel and transmit power in WNs. These strategies are applied to the decentralized SR problem, where independent WNs learn selfishly, based on their own experienced performance.} On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, and shed light on the future of decentralized approaches. Notably, we observe that even though players act selfishly, some of the algorithms learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments. \textcolor{blue}{Finally, the adversarial setting in WNs is studied under two learning implementations: namely \textit{concurrent} and \textit{sequential}. Both procedures rule the operation followed by learners (based on the proposed action-selection strategies). In particular, WNs select an action at the same time for the concurrent approach. In contrast, an ordered action-selection procedure is followed for the sequential case. We study the performance of the aforementioned techniques in terms of convergence speed, average throughput and variability.} The main contributions of this work are summarized below:
	\begin{itemize}		
		\item We devise the feasibility of applying MAB algorithms as defined in the online learning literature to solve the resource allocation problem in WNs.
		\item We study the impact of different parameters intrinsic to the action-selection strategies considered (e.g., exploration coefficients, learning rates) on network performance. \textcolor{blue}{In addition, we analyze the implications derived from the application of different learning procedures, referred to as \textit{concurrent} and \textit{sequential}, which rule the moment at which WNs act.}
		\item \textcolor{blue}{We show the impact of learning concurrently and sequentially. In particular, the former leads to a high throughput variability experienced by WNs, which is significantly reduced by the sequential approach.} \textcolor{blue}{Accordingly, we envision the utilization of sequential approaches to achieve decentralized learning in adversarial wireless networks.}
		\item Finally, we show that there are algorithms that learn to play collaborative actions even though the WNs act selfishly, which is appealing to practical application in chaotic and dynamic environments. In addition, we shed light on the root causes of this phenomena.  
	\end{itemize}

	The remaining of this document is structured as follows: Section \ref{section:related_work} outlines relevant related work. Section \ref{section:mabs} introduces the proposed learning algorithms and their practical implementation for the resource allocation problem in WNs. Then, Section \ref{section:system_model} presents the simulation scenarios and the considerations taken into account. The simulation results are later presented in Section \ref{section:performance_evaluation}. Finally, Section \ref{section:conclusions} provides the final remarks.

	%%%%%%%%%%%%%%%%%%%%%%%
	%%%  II. RELATED WORK 
	%%%%%%%%%%%%%%%%%%%%%%%
	\section{Related Work}
	\label{section:related_work}

	\textcolor{blue}{Decentralized SR has been considerably studied by the wireless research community. The authors in \cite{argyriou2010collision} propose using relay nodes to re-transmit packets lost as a result of a collision. The relay node is able to decode different signals from the environment and to detect if a collision took place. Then, with the aim of improving the re-transmissions operation, it benefits from the current transmission to forward the decoded packets to their original destinations. Although this method shows performance improvements in dense scenarios where collisions are very likely to occur, its effectiveness is subject to the network topology. Regarding directional transmissions, the authors in \cite{babich2015design} propose two novel access schemes to allow multiple simultaneous transmissions. In particular, nodes' activity information is sensed, which, together with antenna's directionality information, allows to build a new set of channel access rules.}

	\textcolor{blue}{Despite approaches based on directional transmissions and interference cancellation are very powerful and allow to significantly increase SR, they strongly rely on having multiple antennas. Such a requirement is not mandatory for the SR operation based on TPC and CST adjustment. In this work, we focus on the former because tuning the transmit power has a direct impact on the generated interference. This allows to purely study the interactions that occur among nodes implementing decentralized SR. Moreover, we consider DCA to be combined with TPC, so that further potential gains can be achieved.}

	\textcolor{blue}{DCA has been extensively studied from the centralized perspective, especially through techniques based on graph coloring \cite{riihijarvi2005frequency, mishra2005weighted}. Despite these kind of approaches allow to effectively reduce the interference between WNs, a certain degree of communication is required. Regarding decentralized methods, the authors in \cite{akl2007dynamic} propose a very simple approach in which each AP maintains an interference map of their neighbors, so that channel assignment is done through interference minimization. Unfortunately, the interactions among APs in the decentralized setting are not studied. Separately, \cite{chen2007improved} proposes two decentralized approaches that rely on the interference measured at both APs and stations (STAs) to calculate the best frequency channels for dynamic channel allocation. To do so, a WN, in addition to the interference sensed by its associated devices, considers other metrics such as the amount of traffic, so that some coordination is required at the neighbor level (e.g., periodic reporting). The authors in \cite{yue2011cacao} show that the decentralized DCA problem is NP-hard. In addition, they propose a distributed algorithm whereby APs select the best channel according to the observed traffic information (i.e., channel sensing is considered).}

	\textcolor{blue}{In this work we aim to extend the approach in \cite{yue2011cacao} in two ways. \textcolor{blue}{First, we aim to provide a flexible solution based on the performance achieved by a given WN.} Second, we aim to tackle the spatial domain through TPC\textcolor{blue}{, which has been shown to provide large improvements in wireless networks \cite{elbatt2000power}}. \textcolor{blue}{However}, dealing with the spatial dimension leads to unpredictable interactions in terms of interference. \textcolor{blue}{Such a complexity is illustrated} in \cite{tang2014joint}, which performs power control and rate adaptation in subgroups of Wireless Local Area Networks (WLANs). The creation of clusters allows defining independent power levels between devices in the same group, which are useful to avoid asymmetric links. However, to represent all the possible combinations, graphs can become very large, especially in high-density deployments. \textcolor{blue}{When it comes to decentralized mechanisms, we find the work in \cite{gandarillas2014dynamic}, which applies TPC based on real-time channel measurements \cite{gandarillas2014dynamic}}. The proposed mechanism (so called Dynamic Transmission Power Control) is based on a set of triggered thresholds that increase/decrease the transmit power according to the state of the system. The main problem is that thresholds are set empirically (based on simulations), which limits the potential of the mechanisms in front of multiple scenarios.}

	\textcolor{blue}{\textcolor{blue}{As shown by previous research, the optimal decentralized SR in WNs through TPC and DCA is very hard to be derived analytically, mostly because of the adversarial setting and the lack of information at nodes. The existing decentralized solutions barely provide flexibility with respect to the scenario, so that potential use cases are disregarded.} For that, we focus on online learning, and more precisely Multi-Armed Bandits (MABs). The MAB framework allows to reduce the complexity of the SR problem, since detailed information about the scenario is not considered. In contrast, learners gain knowledge on all the adversaries as a whole, thus facing a single environment.} To the best of our knowledge, there is very little related work on applying MAB techniques to the problem of resource allocation in WNs. In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in Long Term Evolution (LTE) networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which is shown to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that an optimal resource allocation exists, i.e., every BS obtains the necessary resources. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.

	More related to the problem proposed here, the authors in \cite{maghsudi2015joint} show a channel selection and power control approach in infrastructureless networks, which is modeled through bandits. In particular, two different strategies are provided to improve the performance of two Device to Device (D2D) users (each one composed by a transmitter and a receiver), which must learn the best channel and transmit power to be selected. Similarly to our problem, users do not have any knowledge on the channel or the other's configuration, so they rely on the experienced performance in order to find the best configuration. An extension of \cite{maghsudi2015joint} is provided by the same authors in \cite{maghsudi2015channel}, which includes a calibrated predictor (referred in the work as \textit{forecaster}) to infer the behavior of the other devices in order to counter act their actions. In each agent, the information of the forecaster is used to choose the highest-rewarding action with a certain probability, while the rest of actions are randomly selected. Henceforth, assuming that all the networks use a given strategy $\mathcal{X}$, fast convergence is ensured. Results show that channel resources are optimally distributed in a very short time frame through a fully decentralized algorithm that does not require any kind of coordination. Both aforementioned works rely on the existence of a unique Nash Equilibrium, which favors convergence. In contrast, in this article we aim to extend Bandits utilization to denser deployments, and, what is more important, to scenarios with limited available resources in which there is not a unique Nash Equilibrium (NE) that allows fast-convergence. Thus, we aim to capture the effects of applying selfish strategies in a decentralized way (i.e., agent $i$ follows a strategy $\mathcal{X}_i$ that does not consider the strategies of the others) and we also provide insight about the importance of past information for learning in dense WNs, which has not been studied before.

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%III. MABs APPLICATION									    %	
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Multi-Armed Bandits for Improving Spatial Reuse in WNs}
	\label{section:mabs}

	In this work, we address the decentralized SR problem through online learning because of the uncertainty generated in an adversarial setting. The practical application of MABs in WNs is detailed next:

	% MABs
	\subsection{The Multi-Armed Bandits Framework}
	In the online learning literature, several MAB settings have been considered such as stochastic bandits \cite{thompson1933likelihood,lai1985asymptotically,auer2002finite}, adversarial bandits \cite{auer1995gambling,auer2002nonstochastic}, restless bandits \cite{whittle1988restless}, contextual bandits \cite{LCLS10} and linear bandits \cite{abe2003reinforcement,APS11}, and numerous exploration-exploitation strategies have been proposed such as \textit{$\varepsilon$-greedy} \cite{sutton1998reinforcement,auer2002finite}, \textit{upper confidence bound} (UCB) \cite{lai1985asymptotically,Agr95,BuKa96,auer2002finite}, \textit{exponential weight algorithm for exploration and exploitation} (EXP3) \cite{auer1995gambling,auer2002finite} and \textit{Thompson sampling} \cite{thompson1933likelihood}. The classical multi-armed bandit problem models a sequential interaction scheme between a learner and an environment. The learner sequentially selects one out of $K$ actions (often called \emph{arms} in this context) and earns some rewards determined by the chosen action and also influenced by the environment. Formally, the problem 
	is defined as a repeated game where the following steps are repeated in each round $t=1,2,\dots,T$:
	\begin{enumerate}
		\item The environment fixes an assignment of rewards $r_{a,t}$ for each action $a\in[K] \stackrel{\text{def}}{=} \left\{1,2,\dots,K\right\}$,
		\item the learner chooses action $a_t\in[K]$,
		\item the learner obtains and observes reward $r_{a_t,t}$.
	\end{enumerate}
	The bandit literature largely focuses on the perspective of the learner with the objective of coming up with learning algorithms that attempt to maximize the sum of the rewards gathered during the whole procedure (either with finite or infinite horizon). As noted above, this problem has been studied under various assumptions made on the environment and the structure of the arms. The most important basic cases are the \emph{stochastic} bandit problem where, for each particular arm $a$, the rewards are i.i.d.~realizations of random variables from a fixed (but unknown) distribution $\nu_a$, and the \emph{non-stochastic} (or \emph{adversarial}) bandit problem where the rewards are chosen arbitrarily by the environment. In both cases, the main challenge for the learner is the \emph{partial observability} of the rewards: the learner only gets to observe the reward associated with the chosen action $a_t$, but never observes the rewards realized for the other actions.

@@

	%------------------------
	% Problem formulation
	%------------------------
	\subsection{Multi-Armed Bandits Formulation for Decentralized Spatial Reuse}

	\textcolor{blue}{We model the decentralized SR problem through adversarial bandits. In such a model, the reward experienced by a given agent (WN) is influenced by the whole action profile, i.e., the configurations used by other competing WNs. From a decentralized perspective, the adversarial setting poses several challenges with respect to the existence of a NE. Ideally, the problem is solved if all the competitors implement a pure strategy\footnote{A pure strategy NE is conformed by a set of strategies and payoffs, so that no player can obtain further benefits by deviating from its strategy.} that allows maximizing a certain performance metric. However, finding such a strategy may not be possible in unplanned deployments, due to the competition among nodes and the scarcity of the available resources. Understanding the implications derived from such an adversarial setting in the absence of a NE is one the main goals of this paper, which, to the best of our knowledge, has been barely considered in the previous literature.}

	\textcolor{blue}{In particular,} we model this adversarial problem as follows. Let arm $a \in \mathcal{A}$ (we denote the size of $\mathcal{A}$ with K) be a configuration \textcolor{blue}{in terms of channel and transmit power (e.g., $a_1$ = \{Channel: 1, TPC: -15 dBm\})}. Let $\Gamma_{i,t}$ be the throughput experienced by $\text{WN}_i$ at time $t$, and $\Gamma_{i}^*$ \textcolor{blue}{the optimal throughput.}\footnote{\textcolor{blue}{The optimal throughput is achieved in case of isolation (i.e., when no interference is experienced in the selected channel).}} We then define the reward $r_{i,t}$ experienced by $\text{WN}_i$ at time $t$ as:
